---
title: "HW5 GGPLOT Revisited"
author: "Q Liang"
date: "2/28/2024"
output:
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    fig_width: 6
    number_sections: yes
    toc_collapsed: yes
    code_folding: hide
    code_download: yes
    smooth_scroll: true
    theme: readable
    fig_height: 4
---

```{=html}
<style type="text/css">

div#TOC li {
    list-style:none;
    background-color:lightgray;
    background-image:none;
    background-repeat:none;
    background-position:0;
    font-family: Arial, Helvetica, sans-serif;
    color: #780c0c;
}

/* mouse over link */
div#TOC a:hover {
  color: red;
}

/* unvisited link */
div#TOC a:link {
  color: blue;
}



h1.title {
  font-size: 24px;
  color: Darkblue;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 {
    font-size: 24px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 {
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { 
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}

/* unvisited link */
a:link {
  color: green;
}

/* visited link */
a:visited {
  color: green;
}

/* mouse over link */
a:hover {
  color: red;
}

/* selected link */
a:active {
  color: yellow;
}

</style>
```
```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
options(repos = list(CRAN="http://cran.rstudio.com/"))
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}

              # Install RCurl package
    

if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
### ggplot and extensions
if (!require("ggplot2")) {
    install.packages("ggplot2")              
    library("ggplot2")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("ggridges")) {
    install.packages("ggridges")              
    library("ggridges")
}
if (!require("graphics")) {
    install.packages("graphics")              
    library("graphics")
}

knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE,
                      comment = NA)
```

\

# Prepare a single data set

1.  Reshape data set: Income Per Person to make a longitudinal data such that the resulting data set has three columns: country, year, and income. 

```{r, fig.align='center'}
#Read income_per_person.
data0 <- read.csv ("/Users/qianshengliang/Dropbox/STA553/Week 5/income_per_person.csv",
                   header=TRUE,sep=",")
#change name.
colnames(data0)[which(names(data0) == "geo")] <- "country"
#names(data0)
#Wide_to_long reshape
data0a <- data0 %>%
 pivot_longer(
  cols = starts_with("X"),
  names_to = "year",
  names_prefix = "X",
  #names_transform = list(year = as.integer),
  values_to = "income",
  values_drop_na = TRUE)
#check
#head(data0a)
```

2.  Do the same for Life Expectancy in Years so that the resulting data set has three columns: country, year, and life expectancy.

```{r, fig.align='center'}
#Read life_expectancy_years.
data1 <- read.csv ("/Users/qianshengliang/Dropbox/STA553/Week 5/life_expectancy_years.csv",
                   header=TRUE,sep=",")
#change name.
#data1[1:2,1:200]
#names(data0)
colnames(data1)[which(names(data1) == "geo")] <- "country"
#names(data0)
#Wide_to_long reshape
data1a <- data1 %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "year",
    names_prefix = "X",
    #names_transform = list(Year = as.integer),
    values_to = "lifeExp",
    values_drop_na = TRUE)
#check
#head(data1a)
```

3.  Merge/join the above two longitudinal data sets to make a new data set, under name LifeExpIncom that has variables: country, year, lifeExp, and income.

```{r, fig.align='center'}
Life0 <- left_join(data0a, data1a, by = c("country","year"))
#Life0
```

4. Merge LifeExpIncom with country region so that the final data set has information about income, life expectancy, and country region. 

```{r, fig.align='center'}
#Read countries_total.
data2 <- read.csv ("/Users/qianshengliang/Dropbox/STA553/Week 5/countries_total.csv",
                   header=TRUE,sep=",")
#data2
#change name
#names(data2)
colnames(data2)[which(names(data2) == "name")] <- "country"
data2a <- data2[c(1,6)]
#head(data2a)
Life1 <- left_join(Life0, data2a, by = "country")
#Life1
```

5. Merge the previous resulting data set with population size so that the final data set has information about income, life expectancy, population size, and country region.

```{r, fig.align='center'}
#Read population_total.
data3 <- read.csv ("/Users/qianshengliang/Dropbox/STA553/Week 5/population_total.csv",
                   header=TRUE,sep=",")
#head(data3)
#change name.
#data3[1:5,1:10]
#names(data3)
colnames(data3)[which(names(data3) == "geo")] <- "country"
#names(data3)
#Wide_to_long reshape
data3a <- data3 %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "year",
    names_prefix = "X",
    #names_transform = list(Year = as.integer),
    values_to = "pop",
    values_drop_na = TRUE)
#check
#head(data3a)
#Final dataset
LifeExpIncom <- left_join(Life1, data3a, by = c("country","year"))
#LifeExpIncom
#write.csv (LifeExpIncom, "/Users/qianshengliang/Dropbox/STA553/Week 5/LifeExpIncom.csv")
DT::datatable(LifeExpIncom, fillContainer = FALSE, options = list(pageLength = 10))
```

This data set has 42267 observations and 6 variables.

Create a subset of the above resulting longitudinal data set that contains only 2000 data - name it data2000.

```{r, fig.align='center'}
#2000data
year.id=which(LifeExpIncom$year=="2000")
data2000 = LifeExpIncom[year.id,]
#data2000
#sum(is.na(data2000))
#write.csv (data2000, "/Users/qianshengliang/Dropbox/STA553/Week 5/data2000.csv")
DT::datatable(data2000, fillContainer = FALSE, options = list(pageLength = 10))
```

# Prepare a graph by ggplot

```{r, fig.align='center'}
#ggplot
cols1 = c("#332288","#117733","#44AA99","#88CCEE","#DDCC77","#CC6677")
ggplot(data2000, aes(x = income, y = lifeExp, 
                 color = region, 
                 size = pop)) +
  geom_point(alpha = 0.5) + scale_color_manual(values = cols1) +
  labs(
    x = "Income",
    y = "Life Expectancy",
    # label for legends
    size = "Population:",
    color = "Region:",
    title = "Association between Income and Life Expectancy",
    subtitle = "This is a partial scatter plot",
    caption = paste("Created on", Sys.Date())) +
  theme_minimal()   # minimal theme
```

The graph shows the relation between income and life expectancy of 193 countries/regions in 2000. The size of the points is proportioned to population and colors represent different continents. From this graph, we can find that:

-  Most countries that have low life expectancy  are in Africa. 
-  Most countries that have high life expectancy are in Europe.
-  When income is under 5,000, life expectancy sharply increases to about 70 years when income increases. 
-  When income is 5,000 to 30,000, life expectancy continues to increase slowly with income. 
-  When income is 30,000 to 100,000, life expectancy slightly decreases with income.
-  For the countries with longest life expectancy, the income is 30,000 to 40,000  and the population is usually under 250 million.








